### 2021-04-10 paper/parlr_paper_1.Rmd
Conflict Resolution and Meeting Notes

## Decouple/Recouple

(This will be revised; I'm not sure if I really believe the "Recouple" step anymore! I'll need look into it more on larger datasets)

One difficulty in record linkage is conducting linkage while taking appropriate dependencies into account. In the context of this paper in which we assume that each file is recorded without duplicates, this means our linkage procedure should respect one-to-one matching. The posterior sampling of $\mathbf{Z}$ outlined above resolves half of this issue; since we only sample one potential match for record $j$ in each Gibbs iteration, it is impossible to match $j$ to two records at once. This in itself is an improvement over the classic Fellegi-Sunter method that labels record pairs as matches or non matches independently of all other pairs. However, our posterior sampling of $\mathbf{Z}$ does allow for the possibility that $Z_j = i = Z_{j'}$ for some $j' \neq j$, which cleary violates the assumption that $\mathbf{X}_B$ is recorded without duplicates. Sadinle 2017 resolves this issue by adapting the set of candidate matches for every $Z_j$, but we forgo this rigidity and instead "decouple" the linkage vector $\mathbf{Z}$ for the computational advantantages of sampling $Z_j$ independently. In our formulation, such illegal matches occasionally occur within individual Gibbs iterations due to simple randomness, but they do not occur consistently enough between the same records to measurably influence the eventual Bayes estimate of the linkage structure.

However, if we want to ensure the theoretic purity of each MCMC sample of the linkage structure, we can "recouple" $\mathbf{Z}$ by simply resolving these illegal matchings. Explictly, define sets $R(i) := \{j | Z_j = i\}$, and define all sets such that $|R(i)| > 1$ as *resolution sets*.  Within each set, resample the candidate matches according to their Fellegi-Sunter weights. We show this step to be computationally inexpensive and to be a reasonable replication of more organic one-to-one matching. This procedure would be most useful in cases where there are many "near matches" across files, but in most cases, we find this this step to be unnecessary.

@@

Same as Sadinle.

## Efficient Computation

Adopting some from Enamorado 2019, some steps my own. very promising, stay tuned!

\newpage

## Proposed Linkage Cluster Model

(Still thinking through best notation and terminology)

The primary contribution of this paper is the establishment of linkage clusters, which allow the reliability parameters $\mathbf{m}^f$, the discernment parameters $\mathbf{u}^f$, and also the linkage probability parameter $\lambda$ to vary according to certain attributes in the record pairs. We shall see that this set up upweights linkage within clusters and downweights linkage across clusters, thus using attribute information in a way not yet explored in the literture.

Partition files $A$ and $B$ into *record clusters* $\{A_1, \ldots, A_k\}$ and $\{B_1, \ldots, B_k\}$ according to some attribute in the data. We assign the record pair  $(i,j)$ to a *linkage cluster* $S_k \in \{S_1, \ldots, S_K\}$ only when both records belong the corresponding record clusters; otherwise we assign it to $S_0$. For example
$$i \in A_1, j \in B_1 \implies \gamma_{ij} \in S_1$$
$$i \in A_1, j \in B_2 \implies \gamma_{ij} \in S_0$$
Cluster assignments at the record pair level are necessary for the updating the $\mathbf{m}$ and $\mathbf{u}$ parameters. We denote the reliability and discernment parameters assoicated with each field for each cluster as $\mathbf{m}^{fk}$ and $\mathbf{u}^{fk}$ respectively. To improve readability, we modify the definition of $\mathbf{m}_{ij}$ to incorporate the linkage cluster assignment. Explicitly, 
$$\mathbf{m}_{ij} = \prod_{f=1}^{F}\prod_{l=1}^{L_f}\prod_{k=0}^{K}\left[ \left(m^{flk}\right)^{\mathbf{1}_{\gamma_{ij} \in S_k}}\right]^{\mathbf{1}_{\gamma_{ij}^f = l}}$$
with an analogous definition for $\mathbf{u}_{ij}$.

Additionally, define the vector of linkage probabilities $\Lambda_k = (\lambda_{k1}, \lambda_{k2}, \lambda_{k3})$, where $\lambda_{k1} = P(z_{j\in S_k} \in S_k)$ is the probability that a record in $B_k$ has a match within its cluster in $A_k$, $\lambda_{k2} = P(z_{j \in S_k} \in S_{k'})$ is the probability that such a record $B_k$ has a match outside of its cluster, in  $\mathbf{X}_{A_{k'}}$ for some $k' \neq k$, and $\lambda_{k3} = P(z_{j\in S_k} = n_A + 1)$ is the probability of a non-match. This is a multinomial distribution where we necessarily have $\lambda_{k3} = 1 - \lambda_{k1} - \lambda_{k2}$. These parameters use linkage cluster assignment at the record level.

Our new model specification becomes

$$\mathbf{m^{fk}} \sim \text{Dirichlet}(\alpha^{f1k}, \ldots, \alpha^{fL_fk})$$

$$\mathbf{u^{fk}} \sim \text{Dirichlet}(\beta^{f1k}, \ldots, \beta^{fL_fk})$$
$$Z_{j \in S_k} | \Lambda_k =
\begin{cases} 
    \frac{1}{n_{A_k}}\lambda_{k1}  & z_j \in S_k; \\
    \frac{1}{n_A - n_{A_k}}\lambda_{k2}  & z_j \in S_{k'}; \\
     \lambda_{k3} & z_j = n_A + 1 \\
\end{cases}$$

$$\Lambda \sim \text{Dirichlet}(\Lambda_0)$$
where $\Lambda_0 = (\Lambda_{k1}, \Lambda_{k2}, \Lambda_{k3}).$

**THIS WOULD GO IN APPENDIX**: Through similar derivations, this yields full conditionals:
The $\mathbf{m}$ and $\mathbf{u}$ parameters are updated through standard multinomial-dirichlet mechanics. Thus we have 
$$\mathbf{m^{fk}}|\mathbf{Z}, \Gamma \sim \text{Dirichlet}(\alpha^{f1k}(\mathbf{Z}), \ldots, \alpha^{fL_fk}(\mathbf{Z}))$$

$$\mathbf{u^{fk}}|\mathbf{Z}, \Gamma \sim \text{Dirichlet}(\beta^{f1k}(\mathbf{Z}), \ldots, \beta^{fL_fk}(\mathbf{Z}))$$
where $$\alpha_{flk}(\mathbf{Z})= \sum_{i,j, k} \mathbf{1}_{obs(\gamma_{ij}^f)}\mathbf{1}_{\gamma_{ij}^f = l} \mathbf{1}_{z_j = i}\mathbf{1}_{\gamma_{ij} \in S_k}$$

$$\text{and } \beta_{flk}(\mathbf{Z})= \sum_{i,j, k} \mathbf{1}_{obs(\gamma_{ij}^f)}\mathbf{1}_{\gamma_{ij}^f = l} \mathbf{1}_{z_j \neq i}\mathbf{1}_{\gamma_{ij} \in S_k}$$

We also have straightforward posterior distritutions for the linkage probability vector $\Lambda_k$:

$$\Lambda_k | \mathbf{Z} \sim \text{Dirichlet}(\mathbf{D_k} + \Lambda_0)$$

where $\mathbf{D}_k = (D_{k1}, D_{k2}, D_{k3})$ with 
$$D_{k1} = \sum_{j \in B_k} \mathbf{1}_{z_j \in A_k}$$ $$D_{k2} = \sum_{j \in B_k} \mathbf{1}_{z_j \in A_{k'}}$$ $$\text{and } D_{k3} = \sum_{j \in B_k} \mathbf{1}_{z_j = n_A + 1}$$

Our full conditional for $Z_j$ follows a familiar form:

$$Z_{j \in S_k}^{(s+1)} | \mathbf{m}^k, \mathbf{u}^k, \Gamma, \Lambda_k =
\begin{cases} 
    \frac{\lambda_{k1}}{n_{A_k}} w_{ij} & z_j \in S_k; \\
    \frac{\lambda_{k2}}{n_A - n_{A_k}} w_{ij}  & z_j \in S_{k'}; \\
     \lambda_{k3} & z_j = n_A + 1 \\
\end{cases}$$

And after integrating over $\Lambda_k$ and rearranging terms,

$$Z_{j \in S_k}^{(s+1)} | \mathbf{m}^k, \mathbf{u}^k, \Gamma,  =
\begin{cases} 
    \frac{D_{k1} + \Lambda_{01}}{n_{A_k}(n_B - D_{k1} - D_{k2} + \Lambda_{03})} w_{ij} & z_j \in S_k; \\
    \frac{D_{k2} + \Lambda_{02}}{(n_A - n_{A_k})(n_B - D_{k1} - D_{k2} + \Lambda_{03})} w_{ij}  & z_j \in S_{k'}; \\
     1 & z_j = n_A + 1 \\
\end{cases}$$

>>>>

@@

Our Gibbs sampler provides posterior samples of $\mathbf{Z}$ which we use to make our final decisions about the linkage structure. In the case where we false matches and missed matches contribute the same loss, we declare $(i,j)$ to be a match whenever $P(Z_j = i) > \frac{1}{2}$ according to these posterior samples. We can create more elaborate decisions by attributing different loss values to different kinds of error, and allowing pairings with middling posterior probabilities to be left unlabled by the algorithm so that the modeller can address those manually. For further discussion, see Sadinle 2017.

# Efficient Computation

Broadly speakings, we increase our computational efficiency by tecognizing that record pairs contribute to posterior calculations only through the agreement pattern of the $\gamma_{ij}$ vector. Let $H$ be the set of unique agreement patterns in the data, let $P$ denote the total number of unique agreement patterns present in $\Gamma$, and note that $P$ is bounded above by $\prod_{f=1}^F L_f$. We index these agreement patterns by $p \in \{1, \ldots, P\}$, and say $\gamma_{ij} \in h_p$ when the $(i,j)$ pair exhibits the $p^{th}$ agreement pattern. Wherever possible, we conduct calculcations over these $P$ agreement patterns rather than the $n_A \times n_B$ record pairs.

## Representation of the Data

The classic Fellegi Sunter represents the $\gamma_{ij}$ comparison vector as vector of length $F$, with each component $\gamma_{ij}^f$ taking on values in  $\{0, \ldots, L_f - 1\}$. To ease computations, we instead represent the comparison as a concatenation of $F$ many binary indicators of lengths $L_f$. For example, if $L_1 = L_2 = 2$ and $L_3 = 3$, then $\gamma_{ij} = (1, 0, 2)$ under the classical framework becomes $\gamma_{ij} = (0, 1, 1, 0, 0, 0, 1)$ under our framework. This is a bijective transformive that does not change the meaning of the data, but this representation will ease calculations and posterior updates.

## Hashing

In the classic Fellegi Sunter framework, $\Gamma$ is a $n_A n_B \times F$ matrix, with each row providing the comparison vector for a different $(i,j)$ pair. We however do not store these comparison vectors themselves, but instead only the a hashed value indicating the unique agreement pattern. We store this information in a nested list. For each $p$, we also calculate $|h_p|$ total instances of agreement pattern $h_p$ in the entire $\Gamma$ matrix, and then for each $j$, we also calculate $|h_p|_j$ the total instances agreement pattern for each $\Gamma_{.j}$

## Updating m and u

After receiving matching statuses from $\mathbf{Z}$, the Sadinle method calculates $\alpha_{fl}(\mathbf{Z})$ and $\beta_{fl}(\mathbf{Z})$ for each field and level. This constitutes $2 \times \sum L_f$ many summations over $n_A \times n_B$ quantities, and becomes computionally burdensome with large data. In contrast, we recognize that each unique agreement pattern contributes to the posterior $\alpha(\mathbf{Z})$ and $\beta(\mathbf{Z})$ vectors in the same way. In fact, if we denote $\mathbf{Z}(h_p)$ to be the number of record pairs with agreement pattern $h_p$, then the contribution of pairs of pattern $h_p$ to $\alpha(\mathbf{Z})$ is simply $\mathbf{Z}(h_p) \times h_p$. Thus our posterior update is simply $\alpha(\mathbf{Z}) = \alpha_0 + \sum_{p=1}^P \mathbf{Z}(h_p) \times h_p$. Then $\beta(\mathbf{Z}) = \beta_0 + \sum_{p=1}^P |h_p| - \left(\mathbf{Z}(h_p) \times h_p \right)$

## Updating Z

Sadinle uses the following full conditional for $Z_j$ in order to strictly enforce one-to-one matching.

The indicator removes previously matched records from the set of candidate records, and $L_{-j}$ updates iteratively after each matching decision, $n_B$ many times per Gibbs iteration. This process is *inherently serial* and becomes computationally burdensome for large data. By weakening the one-to-one requirement, our full conditional for $Z$ does not depend on the rest of the $\mathbf{Z}_{-j}$ vector, and thus can be computed in parallel. More importantly, since only the aggrement pattern of $Z_j$ is used for calculations within the Gibbs sampler, and not the particular record label, we can conduct this sampling only at the level of the unique agreement patterns. This boosts computation time far greater than parallelization.

To do this

##
